ABM simulating short-term mating behaviors

### 2022-05-13 submission/manuscript.Rmd
minor language edits of the manuscript, tracked the model directory, close #18

---
output:
  pdf_document:
    keep_tex: true
  word_document: true
geometry: margin=1.0in
fontsize: 11pt
numbersections: true
bibliography: reference.bib
csl: evolution-and-human-behavior.csl
header-includes:
  - \newcommand{\bcenter}{\begin{center}}
  - \newcommand{\ecenter}{\end{center}}
  - \newcommand{\btitlepage}{\begin{titlepage}}
  - \newcommand{\etitlepage}{\end{titlepage}}
  - \usepackage{setspace}\doublespacing
  - \usepackage{helvet}
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{booktabs}
  - \usepackage[font=small,labelfont=bf]{caption}
---

@@

# **This is an awesome title** {-}

@@

Yurun Ying^1^, Pekka Santtila^1^^,^^2^^,^^$\dagger$^

@@

^2^ NYU-ECNU Institute for Social Development at NYU Shanghai

@@

^$\dagger$^ To whom correspondence should be addressed:

@@

Gender differences in short-term mating behaviors have been a well-documented phenomenon in human sexuality research. Existing studies usually conflate gender differences in preferences for short-term mating with differences in sexual behaviors, which is theoretically dubious. Using an agent-based model, the present study investigated whether men and women's different mating preferences resulted in any gender differences in short-term mating behaviors, and if they did, under what circumstances. The results suggested that heterosexual men and women had the same average number of short-term mating experiences and short-term mates even when gender differences in preferences fort-term mating existed. Gender differences in mating behaviors only emerged when heterosexual men and women in the mating pool were considered. Moreover, gay men had a higher average number of both outcomes than lesbian women and than heterosexual men under such a circumstance. These results suggest that theoretically speaking, gender differences in short-term mating behaviors only occurred among particular populations, or when men's preferences for short-term mating are not constrained by those of women. Suggestions for future research in human mating psychology and behaviors were provided.

@@

Intuitive as this line of reasoning is, there are good reasons to doubt its soundness. This is because any heterosexual sexual encounter involves both a man and a woman (the frequency of encounters involving more than two persons is negligible, [e.g., @herbenick_sexual_2017]. A new short-term mating experience for a man is, therefore, also a new experience for a woman, and the same logic applies to counting a new short-term mate. As a result, the psychological differences in short-term mating preferences may not result in behavioral ones in the heterosexual case [@archer_reality_2019].

The present study took gender differences in mating preferences as an assumption, which was proposed by sexual strategy theory [@buss_sexual_1993] and repeatedly found support by empirical studies [e.g., @schmitt_universal_2003; @walter_sex_2020]. Using an agent-based model, we investigated whether men and women's different mating preferences resulted in any gender differences in short-term mating behaviors, specifically, in the number of short-term mating experiences and short-term mates, and if they did, under what circumstances.

@@

Gender differences in mating strategies have been studied in the light of sexual strategy theory [@buss_sexual_1993]. It posits that since the minimum investment that men devote to their offspring (contribution of sperm through one sexual act) is lower than that of women (gestation, labor, and lactation), men tend to have short-term mating as a larger part of their mating strategy than women do. This is because women's high minimum investment leads to higher opportunity cost (vs. men) if they have sex with a suboptimal partner and give birth to offspring with a low survival chance. As a result, men may have a greater interest in short-term mating, desire a larger number of short-term mates in a given period, and be less selective with respect to accepting a potential mate [@buss_sexual_1993].

@@

As for mating standards, men are more likely to accept someone as a potential short-term mate. For example, studies using U.S. college samples found that the minimum percentile ranks that men found acceptable for a potential short-term sexual partner were lower than for women both overall and on individual traits (e.g., social status, attractiveness) [@kenrick_evolution_1990; @regan_minimum_1998]. A study also found that when presented with identical descriptions of potential mates, men on average rated them as more desirable than women did [@wiederman_evolution_1998].

Some evidence shows that sexual strategy theory also predicts mating preferences among gay men and lesbian women. A survey study using a community sample from the U.S. found that gay men were more interested in short-term mating than lesbian women and that this gender difference was comparable to that existing among heterosexual individuals [@bailey_effects_1994]. An indirect piece of evidence on differential standards for short-term mates comes from a recent study finding that significantly more gay men than lesbian women reported to have accepted a casual sexual offer from a same-gender person [@matsick_gender_2021]. Past studies suggest that the gender difference in the acceptance rate of casual sexual offers may originate from men and women’s differential standards for short-term mates [@conley_women_2011; @hald_receptivity_2010]. Thus, as a postulation, gay men and lesbian women may also have different standards for short-term mates.

## Constraints on men's preferences for short-term mating
Since heterosexual sex involves both a man and a woman in most cases, men's interest in short-term mating can be constrained by women's preference for long-term relationships [@archer_reality_2019; @symons_evolution_1979]. This is because men’s short-term mating preferences can only translate into behaviors when there are women willing to have sex with them. When short-term mating occurs, the number of short-term mating encounters and short-term mates is essentially the same for men and women (although this does not necessarily mean that the total number of men and women who have ever had short-term mating must be equal). Therefore, we would expect to observe no gender difference in short-term mating behaviors among heterosexual individuals even when there were gender differences in mating preferences.

As a comparison, in the cases of gay men and lesbian women, men’s preferences are not constrained by women’s, but only by those of other men, who, arguably, have more similar preferences. This would allow for a more direct behavioral expression of men’s mating preferences. The notion that gay men have less restricted preferences was borne out in the proportion of individuals who have engaged in extradyadic sex [@peplau_sexuality_2004]. A study showed that the proportions of heterosexual men and women who have engaged in extradyadic sex were 26% and 21%, respectively, while the statistics among gay men and lesbian women were 82% and 28%, respectively [@peplau_sexuality_2004]. Therefore, we would theoretically expect to observe gender differences in short-term mating behaviors among gay men and lesbians. Moreover, we would also expect to find that gay men, as compared to heterosexual men, engaged in more short-term mating behaviors due to the lessened constraint on their mating preferences.

@@

By using a spatial agent-based model, the present study investigated whether men and women's different mating preferences resulted in any gender differences in short-term mating behaviors, and if they did, under what circumstances. The interest in short-term mating was modeled as an individual’s likelihood of deciding to have short-term mating at each time step. The standards for short-term mates were modeled as the minimum desirability of a potential mate with whom an individual was willing to have sex. Short-term mating behaviors were operationally defined as the number of short-term mating experiences and the number of past short-term mates.

We modeled this process among both heterosexual individuals and gay men and lesbian women to examine whether constraints set by the opposite sex’s preferences would make a difference in short-term mating behaviors. Individuals’ sexual orientation was conceptualized in terms of behaviors only in our model. Heterosexual men had short-term mating or formed a long-term relationship with women, while gay men only with other men, and vice versa for women.

@@

The model measured two outcomes: (1) the number of short-term mating experiences of men and women; (2) the number of short-term mates of men and women. Additionally, we had also documented the number of men and women in the mating pool (i.e., who had ever engaged in short-term mating); The average number of short-term mating experiences and short-term mates were calculated by taking the average among the whole population of men/women (*N~m~* = *N~w~* = 150) and among those who had engaged in short-term mating.

See the Supplemental Materials for the overview, design concepts, and details (ODD) protocol of the model, which includes detailed scheduling and parameterization. The model can be downloaded from the Github online repository [link masked for peer review]. The model was programmed in Netlogo 6.2.1 [@wilensky_1999].

@@

Using the agent-based model, the present study conducted two 2 (gender difference vs. no difference in short-term mating likelihood) x 2 (gender difference vs. no difference in mating standard) experiments. Experiment 1 was run among heterosexual agents who only engaged in short-term or long-term relationships with agents of the opposite gender. Experiment 2 was run among gay men and lesbian women who only engaged in short-term or long-term relationships with agents of the same gender.

Gender differences were modeled in agents’ likelihood of engaging in short-term mating and the standard for short-term mates. The values of these parameters were set based on empirical findings in human mating psychology. We compared the conditions where there were gender differences in mating preferences with counterfactual conditions where gender differences were missing. When there was a gender difference in short-term mating likelihood, men had a 40% likelihood of deciding to engage in short-term mating upon meeting a potential mate, while women had a likelihood of 25% [@buss_sexual_1993; @schmitt_are_2001; @schmitt_universal_2003]. When there was no gender difference in short-term mating likelihood, both women and men had a likelihood of 25%. When there was a gender difference in mating standards, men had a mating standard of 3 (the minimum mate value of a potential partner with whom an agent is willing to have short-term mating, highest possible value = 10), while women had a standard of 5 [@kenrick_evolution_1990; @regan_minimum_1998]. When there was no gender difference in mating standards, both men and women had a standard of 5.

@@

The likelihood of engaging in short-term mating and the standard for short-term mates were initialized either the same or differently for men and women, depending on the experimental conditions.

@@

#### Gay men and lesbian women procedures.
At each time step, men and women moved and decided on short-term mating as the agents did in the heterosexual procedures. Half of the agents (an equal number of men and women) were also given the initiator status. The initiators checked to see whether there were other agents at the same location. The rest of the procedures were identical to the heterosexual procedures, except that the agents only chose those with the same gender as themselves as potential short-term mates or long-term partners.

## Simulations
The model was run for 1,000 time steps in each simulation. We ran 10,000 simulations for each experiment, and 2,500 simulations for each condition. We controlled for initializing random seeds in the model runs. All simulations were run using Netlogo 6.2.1 [@wilensky_1999].

@@

## Gender differences in short-term mating behaviors in each experiment

```{r sec1.1}
m_exp_all.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_exp_all.mean)
m_exp_all.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_exp_all.sd)
f_exp_all.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_exp_all.mean)
f_exp_all.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_exp_all.sd)
exp_all.d_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_all.d) %>% abs()

m_partner_all.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_partner_all.mean)
m_partner_all.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_partner_all.sd)
f_partner_all.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_partner_all.mean)
f_partner_all.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_partner_all.sd)
partner_all.d_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(partner_all.d) %>% abs()
```

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=80mm]{figures/fig1_men_vs_women.pdf}
  \caption{\textbf{Short-term mating behaviors of men and women after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables separately for heterosexual individuals and gay men and lesbian women. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. Statistics were calculated using the full population of men and women in the model (\textit{N\textsubscript{m}} = \textit{N\textsubscript{w}} = 150).}
\end{figure}

```{r sec1.2}
m_exp_all.mean_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(m_exp_all.mean)
m_exp_all.sd_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(m_exp_all.sd)
f_exp_all.mean_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(f_exp_all.mean)
f_exp_all.sd_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(f_exp_all.sd)
exp.all.parameter_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(exp_all.parameter)
exp.all.statistic_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(exp_all.statistic)
exp.all.p.value_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(exp_all.p.value)
exp_all.d_111 <-
results %>% filter(hls == "1_1_1") %>% pull(exp_all.d) %>% abs()

m_partner_all.mean_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(m_partner_all.mean)
m_partner_all.sd_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(m_partner_all.sd)
f_partner_all.mean_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(f_partner_all.mean)
f_partner_all.sd_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(f_partner_all.sd)
partner.all.parameter_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(partner_all.parameter)
partner.all.statistic_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(partner_all.statistic)
partner.all.p.value_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(partner_all.p.value)  
partner_all.d_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(partner_all.d) %>% abs()
```

@@

```{r sec1.3}
m_exp_ip.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_exp_ip.mean)
m_exp_ip.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_exp_ip.sd)
f_exp_ip.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_exp_ip.mean)
f_exp_ip.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_exp_ip.sd)
exp_ip.parameter_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.parameter)
exp_ip.statistic_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.statistic)
exp_ip.p.value_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.p.value)
exp_ip.d_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.d) %>% abs()

m_partner_ip.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_partner_ip.mean)
m_partner_ip.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_partner_ip.sd)
f_partner_ip.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_partner_ip.mean)
f_partner_ip.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_partner_ip.sd)
partner_ip.parameter_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(partner_ip.parameter)
partner_ip.statistic_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(partner_ip.statistic)
partner_ip.p.value_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(partner_ip.p.value)
partner_ip.d_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(partner_ip.d) %>% abs()
```

In Experiment 1, however, when the means of the outcome variables were calculated among men and women who were in the mating pool (i.e., those with a non-zero number of short-term mating experiences) **(the number of individuals in the mating pool: M~m~, SD~m~; M~w~, SD~w~)**, there were gender differences in short-term mating behaviors when there were gender differences in mating preferences. The average number of short-term mating experiences was higher among heterosexual men (*M* = `r m_exp_ip.mean_011`, *SD* = `r m_exp_ip.sd_011`) than among heterosexual women (*M* = `r f_exp_ip.mean_011`, *SD* = `r f_exp_ip.sd_011`), *t*(`r exp_ip.parameter_011`) = `r exp_ip.statistic_011`, *p* `r exp_ip.p.value_011`, Cohen's *d* = `r exp_ip.d_011`. Likewise, the average number of short-term mates was higher among heterosexual men (*M* = `r m_partner_ip.mean_011`, *SD* = `r m_partner_ip.sd_011`) than among heterosexual women (*M* = `r f_partner_ip.mean_011`, *SD* = `r f_partner_ip.sd_011`), *t*(`r partner_ip.parameter_011`) = `r partner_ip.statistic_011`, *p* `r partner_ip.p.value_011`, Cohen's *d* = `r partner_ip.d_011`.

## Comparing short-term mating behaviors of heterosexual and gay men

```{r sec2}
m_exp_all.mean_0_11 <-
  results_hg %>% filter(ls == "1_1") %>% pull(m_exp_all.mean_0)
m_exp_all.sd_0_11 <-
  results_hg %>% filter(ls == "1_1") %>% pull(m_exp_all.sd_0)
m_exp_all.mean_1_11 <-
  results_hg %>% filter(ls == "1_1") %>% pull(m_exp_all.mean_1)
m_exp_all.sd_1_11 <-
  results_hg %>% filter(ls == "1_1") %>% pull(m_exp_all.sd_1)
exp.all.parameter_hg <-
  results_hg %>% filter(ls == "1_1") %>% pull(m_exp_all.parameter)
exp.all.statistic_hg <-
  results_hg %>% filter(ls == "1_1") %>% pull(m_exp_all.statistic)
exp.all.p.value_hg <-
  results_hg %>% filter(ls == "1_1") %>% pull(m_exp_all.p.value)
exp_all.d_hg <-
  results_hg %>% filter(ls == "1_1") %>% pull(m_exp_all.d) %>% abs()

>>>>

---
output:
  pdf_document:
    keep_tex: true
geometry: margin=1.0in
fontsize: 11pt
numbersections: true
bibliography: reference.bib
csl: evolution-and-human-behavior.csl
header-includes:
  - \newcommand{\bcenter}{\begin{center}}
  - \newcommand{\ecenter}{\end{center}}
  - \newcommand{\btitlepage}{\begin{titlepage}}
  - \newcommand{\etitlepage}{\end{titlepage}}
  - \usepackage{setspace}\doublespacing
  - \usepackage{helvet}
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{booktabs}
  - \usepackage[font=small,labelfont=bf]{caption}
---

@@

# An Agent-based Model of Gender Differences in Short-term Mating Behaviors as a Result of Mating Preferences {-}

@@

Yurun Ying^1^, Jan Antfolk^2^, Pekka Santtila^1^^,^^3^^,^^$\dagger$^

@@

^2^ xxxx

^3^ NYU-ECNU Institute for Social Development at NYU Shanghai

@@

^$\dagger$^ To whom correspondence should be addressed: [pekka.santtila@nyu.edu](pekka.santtila@nyu.edu)

@@

Gender differences in short-term mating behaviors have been a well-documented phenomenon in human sexuality research. Existing studies usually conflate gender differences in mating preferences with differences in sexual behaviors, which is theoretically dubious. Using an agent-based model, the present study investigated the circumstances under which men and women’s differential preferences for short-term mating would result in gender differences in short-term mating behaviors. The results from the model suggested that heterosexual men and women had the same average number of short-term mating experiences and short-term mates even when men had stronger preferences for short-term mating. Gender differences in mating behaviors only emerged when heterosexual men and women in the mating pool (i.e., those with a non-zero number of short-term mating experiences) were considered. Moreover, gay men had a higher average number of both outcomes than lesbian women as well as than heterosexual men when men had stronger preferences for short-term mating. These results suggest that theoretically speaking, even when gender differences in mating preferences existed, those in short-term mating behaviors only occurred among particular populations, or when men’s preferences for short-term mating are not constrained by those of women. Suggestions for future research in human mating psychology and behaviors were provided.

@@

Intuitive as this line of reasoning is, there are good reasons to doubt its soundness. This is because any heterosexual sexual encounter involves both a man and a woman [the frequency of encounters involving more than two persons is negligible, e.g., @herbenick_sexual_2017]. A new short-term mating experience for a man is, therefore, also a new experience for a woman, and the same logic applies to counting a new short-term mate. As a result, the psychological differences in short-term mating preferences may not result in behavioral ones in the heterosexual case [@archer_reality_2019].

The present study took gender differences in mating preferences as a presumption, which was conceptualized by sexual strategy theory [@buss_sexual_1993] and repeatedly found support by empirical studies [e.g., @schmitt_universal_2003; @walter_sex_2020]. Using an agent-based model, we investigated whether men and women's differential preferences for short-term mating resulted in any gender differences in short-term mating behaviors, specifically, in the number of short-term mating experiences and short-term mates, and if they did, under what circumstances.

@@

Gender differences in mating preferences have been studied in the light of sexual strategy theory [@buss_sexual_1993]. It posits that since the minimum investment that men devote to their offspring (contribution of sperm through one sexual act) is lower than that of women (gestation, labor, and lactation), men tend to have short-term mating as a larger part of their mating strategy than women do. This is because women's high minimum investment leads to higher opportunity cost (vs. men) if they have sex with a suboptimal partner and give birth to offspring with a low survival chance. As a result, men may have a greater interest in short-term mating, desire a larger number of short-term mates in a given period, and be less selective with respect to accepting a potential mate [@buss_sexual_1993].

@@

As for mating standards, men are less selective in accepting someone as a potential short-term mate. For example, studies using U.S. college samples found that the minimum percentile ranks that men found acceptable for a potential short-term sexual partner were lower than those accepted by women, both in terms of overall desirability and individual traits (e.g., social status, attractiveness) [@kenrick_evolution_1990; @regan_minimum_1998]. A study also found that when presented with identical descriptions of potential mates, men on average rated them as more desirable than women did [@wiederman_evolution_1998].

Some evidence shows that gender differences in mating preferences also existed between gay men and lesbian women. A survey study using a community sample from the U.S. found that gay men were more interested in short-term mating than lesbian women were, and that this gender difference was comparable to that existing among heterosexual individuals [@bailey_effects_1994]. An indirect piece of evidence on differential standards for short-term mates comes from a recent study finding that significantly more gay men than lesbian women reported to have accepted a casual sexual offer from a same-gender person [@matsick_gender_2021]. Past studies suggest that the gender difference in the acceptance rate of casual sexual offers may originate from men and women’s differential standards for short-term mates [@conley_women_2011; @hald_receptivity_2010]. Thus, as a postulation, gay men may also have lower standards than lesbian women for short-term mates.

## Constraints on men's preferences for short-term mating
Since heterosexual sex involves both a man and a woman in most cases, men's interest in short-term mating can be constrained by women's preference for long-term relationships [@archer_reality_2019; @symons_evolution_1979]. This is because men’s short-term mating preferences can only translate into behaviors when there are women willing to have sex with them. When a new short-term mating encounter occurs, it counts towards men’s total number of short-term mating encounters as well as towards women’s. Similarly, a new short-term mate counts towards both men and women’s total number of short-term mates. Therefore, the total number of short-term mating experiences and short-term mates must be equal between heterosexual men and women on a population level. However, it is important to note that although heterosexual men’s total number of short-term mating is equal to that of women, this does not necessarily mean that the total number of men and women who have ever had short-term mating must be equal. It is possible that there is a smaller proportion of men (vs. the proportion of women) who contribute to the total number of short-term mating. Following this line of reasoning, we would expect to observe no gender difference in short-term mating behaviors among heterosexual individuals even when there were gender differences in mating preferences.

As a comparison, in the cases of gay men and lesbian women, men’s preferences are not constrained by women’s, but only by those of other men, who, arguably, have more similar preferences. This would allow for a more direct behavioral expression of men’s mating preferences. The notion that gay men have less restricted preferences was borne out in the proportion of individuals who have engaged in extradyadic sex [@peplau_sexuality_2004]. A study showed that the proportions of heterosexual men and women who have engaged in extradyadic sex were 26% and 21%, respectively, while the statistics among gay men and lesbian women were 82% and 28%, respectively [@peplau_sexuality_2004]. Therefore, we would theoretically expect to observe gender differences in short-term mating behaviors between gay men and lesbian women. Moreover, we would also expect to find that gay men, as compared to heterosexual men, engaged in more short-term mating behaviors due to the lessened constraint on their mating preferences.

@@

By using a spatial agent-based model, the present study investigated whether men and women’s different preferences for short-term mating resulted in any gender differences in short-term mating behaviors, and if they did, under what circumstances. Preferences for short-term mating were operationally defined as the interest in short-term mating and standards for short-term mates [@buss_sexual_1993]. Short-term mating behaviors were operationally defined as the number of short-term mating experiences and the number of past short-term mates [@petersen_meta-analytic_2010].

We modeled this process among both heterosexual individuals and gay men and lesbian women to examine whether constraints set by the opposite sex’s preferences would affect short-term mating behaviors. Individuals’ sexual orientation was conceptualized in terms of behaviors only in our model. Heterosexual men had short-term mating or formed a long-term relationship only with women, while gay men only with other men, and vice versa for women.

@@

The model measured two outcomes: (1) the number of short-term mating experiences of men and women; (2) the number of short-term mates of men and women. Additionally, we also measured the number of men and women in the mating pool (i.e., those with a non-zero number of short-term mating experiences). The average numbers of short-term mating experiences and short-term mates were calculated by taking the average among the whole population of men/women (*N~m~* = *N~w~* = 150) and among those who had engaged in short-term mating.

See the Supplemental Materials for the overview, design concepts, and details (ODD) protocol of the model, which includes detailed scheduling and parameterization. The model can be downloaded from the Open Science Framework [link masked for peer review]. The model was programmed in Netlogo 6.2.1 [@wilensky_1999].

@@

Using the agent-based model, the present study conducted two 2 (gender difference vs. no difference in the interest in short-term mating likelihood) x 2 (gender difference vs. no difference in mating standards) experiments. Experiment 1 was run among heterosexual agents who only engaged in short-term or long-term relationships with agents of the opposite gender. Experiment 2 was run among gay men and lesbian women who only engaged in short-term or long-term relationships with agents of the same gender.

The interest in short-term mating was modeled as an agent’s likelihood of deciding to have short-term mating at each time step. The standards for short-term mates were modeled as the minimum desirability of a potential mate with whom an agent was willing to have sex. The values of the two parameters were set based on empirical findings in human mating psychology. When there was a gender difference in short-term mating likelihood, men had a 40% likelihood of deciding to engage in short-term mating, while women had a likelihood of 25% [@buss_sexual_1993; @schmitt_are_2001; @schmitt_universal_2003]. When there was no gender difference in short-term mating likelihood, both women and men had a likelihood of 25%. When there was a gender difference in mating standards, men had a mating standard of 3 (the highest possible value = 10), while women had a standard of 5 [@kenrick_evolution_1990; @regan_minimum_1998]. When there was no gender difference in mating standards, both men and women had a standard of 5.

@@

The likelihood of engaging in short-term mating and the standard for short-term mates were initialized either identically or differently for men and women, depending on the experimental conditions.

@@

#### Gay men and lesbian women procedures.
At each time step, men and women moved and decided on short-term mating as the agents did in the heterosexual procedures. Half of the agents (75 men and 75 women) were also given the initiator status. The initiators checked to see whether there were other agents at the same location. The rest of the procedures were identical to the heterosexual procedures, except that the agents only chose those with the same gender as potential short-term mates or long-term partners.

## Simulations
The model was run for 1,000 time steps in each simulation. We ran 10,000 simulations for each experiment, including 2,500 simulations for each condition. We controlled for initializing random seeds in the model runs. All simulations were run using Netlogo 6.2.1 [@wilensky_1999].

@@

## Gender differences in short-term mating behaviors

```{r sec1.1}
# the number of experiences, difference*2, heterosexual
m_exp_all.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_exp_all.mean)
m_exp_all.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_exp_all.sd)
f_exp_all.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_exp_all.mean)
f_exp_all.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_exp_all.sd)
exp_all.d_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_all.d) %>% abs()

# the number of partners, difference*2, heterosexual
m_partner_all.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_partner_all.mean)
m_partner_all.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_partner_all.sd)
f_partner_all.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_partner_all.mean)
f_partner_all.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_partner_all.sd)
partner_all.d_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(partner_all.d) %>% abs()
```

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=70mm]{figures/fig1_men_vs_women.pdf}
  \caption{\textbf{Short-term mating behaviors of men and women after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables separately for heterosexual individuals and gay men and lesbian women. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. Statistics were calculated using the full population of men and women in the model (\textit{N\textsubscript{m}} = \textit{N\textsubscript{w}} = 150).}
\end{figure}

```{r sec1.2}
# the number of experiences, difference*2, gay and lesbians
m_exp_all.mean_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(m_exp_all.mean)
m_exp_all.sd_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(m_exp_all.sd)
f_exp_all.mean_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(f_exp_all.mean)
f_exp_all.sd_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(f_exp_all.sd)
exp.all.parameter_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(exp_all.parameter)
exp.all.statistic_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(exp_all.statistic)
exp.all.p.value_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(exp_all.p.value)
exp_all.d_111 <-
results %>% filter(hls == "1_1_1") %>% pull(exp_all.d) %>% abs()

# the number of partners, difference*2, gay and lesbians
m_partner_all.mean_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(m_partner_all.mean)
m_partner_all.sd_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(m_partner_all.sd)
f_partner_all.mean_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(f_partner_all.mean)
f_partner_all.sd_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(f_partner_all.sd)
partner.all.parameter_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(partner_all.parameter)
partner.all.statistic_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(partner_all.statistic)
partner.all.p.value_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(partner_all.p.value)  
partner_all.d_111 <-
  results %>% filter(hls == "1_1_1") %>% pull(partner_all.d) %>% abs()
```

@@

```{r sec1.3}
# the number of experiences, difference*2, heterosexual in the mating pool
m_exp_ip.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_exp_ip.mean)
m_exp_ip.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_exp_ip.sd)
f_exp_ip.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_exp_ip.mean)
f_exp_ip.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_exp_ip.sd)
exp_ip.parameter_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.parameter)
exp_ip.statistic_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.statistic)
exp_ip.p.value_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.p.value)
exp_ip.d_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.d) %>% abs()

# the number of partners, difference*2, heterosexual in the mating pool
m_partner_ip.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_partner_ip.mean)
m_partner_ip.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_partner_ip.sd)
f_partner_ip.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_partner_ip.mean)
f_partner_ip.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_partner_ip.sd)
partner_ip.parameter_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(partner_ip.parameter)
partner_ip.statistic_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(partner_ip.statistic)
partner_ip.p.value_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(partner_ip.p.value)
partner_ip.d_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(partner_ip.d) %>% abs()

# the number of in-pool, difference*2
m_inpool.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_inpool.mean)
m_inpool.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_inpool.sd)
f_inpool.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_inpool.mean)
f_inpool.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_inpool.sd)
```

In Experiment 1, however, when the means of the outcome variables were calculated among men and women in the mating pool (the number of individuals in the mating pool: *M~m~* = `r m_inpool.mean_011`, *SD~m~* = `r m_inpool.sd_011`; *M~w~* = `r f_inpool.mean_011`, *SD~w~* = `r f_inpool.sd_011`), there were gender differences in short-term mating behaviors when men and women had differential preferences for short-term mating. The average number of short-term mating experiences was higher among heterosexual men (*M* = `r m_exp_ip.mean_011`, *SD* = `r m_exp_ip.sd_011`) than among heterosexual women (*M* = `r f_exp_ip.mean_011`, *SD* = `r f_exp_ip.sd_011`), *t*(`r exp_ip.parameter_011`) = `r exp_ip.statistic_011`, *p* `r exp_ip.p.value_011`, Cohen's *d* = `r exp_ip.d_011`. Likewise, the average number of short-term mates was higher among heterosexual men (*M* = `r m_partner_ip.mean_011`, *SD* = `r m_partner_ip.sd_011`) than among heterosexual women (*M* = `r f_partner_ip.mean_011`, *SD* = `r f_partner_ip.sd_011`), *t*(`r partner_ip.parameter_011`) = `r partner_ip.statistic_011`, *p* `r partner_ip.p.value_011`, Cohen's *d* = `r partner_ip.d_011`.

## Comparing heterosexual and gay men
